False Positive Probabilities for all Kepler Objects of Interest: \\ \nvalnew\ newly validated planets and \nfpnew\ likely false positives

### 2016-03-26 document/ms.tex
new hz table

\newcommand{\ntotal}{7470}
\newcommand{\nfail}{427}
\newcommand{\ncalc}{7056}

@@

\newcommand{\nreliable}{2857} %from figure
\newcommand{\nval}{1948} %from figure
\newcommand{\nreliableFP}{419} %from figure
\newcommand{\nvalnew}{1284} %from pie chart
\newcommand{\nfpnew}{428}  % all candidate with FPP < 0.9
\newcommand{\ncurval}{984} % from pie chart
\newcommand{\nhz}{10}

@@

@@

Of the \ntotal\ KOIs in the Q1-Q17 DR24 table at the NExScI Exoplanet
Archive, \vespa\ successfully calculates the FPP for \ncalc.  These
results are listed in \tab{fpp};  \sect{failures} contains detailed
explanations of the failure reasons for the other \nfail\ KOIs.

In order to properly interpret these results, it is necessary to
understand the range of applicability of the \vespa\ calculation.
First of all, this method selects between different specific explanations 
for the transit-like signal, and cannot
comment on whether the signal might be caused by stellar variability
or an instrumental false alarm.  Thus, \vespa\ results on low
signal-to-noise ratio (SNR) candidates that are not clearly
transit-like must be viewed with caution.  This being said, the reason 
for including the artificial 
``boxy'' and ``long'' models in the model selection calculation
 (\sect{methods:fpp}) is to flag signals that do not 
fit well with any of the astrophysical eclipse models---in fact, the ``long''
model is preferred ($>50\5$) by \nlongfp\ KOIs that are already dispositioned
as FALSE POSITIVE.

@@

In the Q1-Q17 DR24 KOI table used as the basis for this work, 980
KOIs were identified as PEMs.  Of these, 187 were not identified 
as false positives by any other method.  Only 15 of these 187 survived 
all the quality cuts described in \sect{results:fpp}. And of these 15, 
only 3 have FPP $< 0.01$  (See \sect{gaidos} for an example of a KOI caused by a 
``column anomaly'' effect that went unidentified by the \kepler\ pipeline but
to which \vespa\ assigned a high FPP.).  Thus, we expect only about 0.3\% (3/980) of 
as-yet unidentified distant-contamination FPs to end
up with FPP $< 0.01$ according to \vespa. As the fraction of false positives
from pixel contamination but unidentified as PEMs among the entire KOI sample is
estimated to be something around 23\% \citep{Coughlin:2014}, we estimate
that there remains a small ($\sim$0.06\%) residual FPP for all KOIs, even 
when the \vespa\-calculated FPP is negligibly tiny.

@@

While the FPP below which to claim planet validation is clearly
an arbitrary choice, there is precedent to using ${\rm FPP} <0.01$ as 
the threshold---\citet{Rowe:2014} validated over 800 multi-planet KOIs
using this number.  Adopting this same threshold, and incorporating
the 0.06\% residual FPP estimated in \sect{PEMs}, we find that 
of the KOIs with reliable \vespa\ FPPs, \nval\ have ${\rm FPP}
<0.0094$, and are thus validated at the 99\% level.  
These KOIs are labeled as such in \tab{fpp}.  These are not all new validations,
however, since a number of them are already CONFIRMED. \Fig{fpppie}
shows a different kind of summary, grouping KOIs by disposition and
splitting up the CANDIDATES according to the \vespa\ results, showing
that \nvalnew\ KOIs are newly validated at the 99\% level.
\Fig{rpcand} shows the radii and periods of the CONFIRMED and 
CANDIDATE KOIs, with the transparency of the points representing
the \vespa-calculated FPP.

@@

In addition to the confident validations, we identify \nfpnew\ KOIs
that currently have CANDIDATE disposition but for which \vespa\
calculates FPP $>$ 0.9; these KOIs are likely to be false positives.
As \Fig{rpcand} shows, many of these newly identified false positives
have large radii---this is again because of the dispositioning
philosophy adopted by the \kepler\ team, which does not use any cut in
transit depth or inferred ``planet'' size to identify FALSE POSITIVES.
To identify these likely false positives from the \vespa\
calculations, we do not require CANDIDATES to obey the same selections
we used to ensure a clean sample for validation.  This is because even
if a signal has characteristics such that a low \vespa\ FPP would not
be sufficient to validate it, a high \vespa\ FPP is still sufficient
to cast doubt on its planetary nature. As a demonstration of the
ability of \vespa\ to identify false positives, \Fig{fppfps} shows the
FPPs for KOIs that are dispositioned FALSE POSITIVE and have MES $>$
10 and $L_{\rm tot} > 10^{-3}$. The vast majority of these have large
FPPs.

@@

In this work, the only false positive scenarios we consider are 
eclisping binary stars (EB, HEB, BEB, and the double-period versions
thereof).  Previous work studying \kepler\ false positive rates
\citep[e.g.][]{Fressin:2013} has also considered the  ``blended transiting
planet'' to be a false positive---i.e.,  a fainter companion
star hosting a transiting planet larger than what would be inferred if
it were transiting the target star.  Because we do not consider such
a scenario to be a false positive, the \vespa\ analysis presented here 
does not quantify its probability.  As a result, we are \textit{not}
able to unambiguously determine the radii of the planets we validate---
all the planet radii listed in \tab{fpp} are based on the assumption
that the planet transits the target star and that the target star is 
unblended.  If the target star is actually a member of an unresolved
binary system (as a significant fraction of KOIs undoubtedly are),
then the true planet radius will be larger (significantly larger 
if transiting a fainter companion).  This was indeed the motivation
for \citet{Fressin:2013} to consider the ``blended transiting planet''
as a false positive; part of the goal of that work was to 
compute the planet occurrence rates in different radius bins.  
However, trying to distinguish between single and binary target
star configurations is beyond the scope of this work; we thus follow
the precedent of \citet[][especially Section 5]{Lissauer:2014} by 
acknowledging the potential for substantial radius
uncertainties among the validated planet sample while nevertheless 
defending the validations themselves as robust.

@@

In this work, we have calculated the astrophysical false positive
probability (FPP) for every \kepler\ object of interest (KOI) in the
Q1-Q17 DR24 table, using the publicly available Python module
\vespa, which implements the procedure introduced in
\citet{Morton:2012}, with improvements in stellar parameter modeling
and the inclusion of new ``double-period'' false positive scenarios.
While the assumptions behind this calculation are not necessarily
valid for every KOI (see \sect{results:fpp}), we have identified
\nvalnew\ KOIs that have reliable FPPs of $< 0.01$, resulting in
validation of their planetary nature at the 99\% confidence level,
more than doubling the number of confirmed \kepler\ exoplanets.  Among
this set of reliable \vespa\ calculations are also \nfpnew\ new likely
false positive KOIs, although we note that some of these may be due to
unidentified or miscorrected transit timing variations.

The reliability of these calculations depends significantly on the
results of \citet{Bryson:KSCI}, which quantify the probability that
the eclipse/transit signal is spatially coincident on the sky with the
presumed target star.  Without confirmation that the transit signal is
not coming from a significantly displaced source, the sky area used as
part of the prior for the false positive scenarios would need to be
significantly larger than the positional uncertainty value assumed in
this work (described in \sect{methods:fpp}).  Additionally, the
blended false positive scenarios \vespa\ considers are assumed to be
fully contained within the photometric aperture; this assumption would
also be broken if the source of the eclipse were significantly
displaced.  We estimate that perhaps 0.1\% of the planets we validate
could be signals coming from significantly displaced sources, similar 
to those identified as period-epoch matches by \citet{Coughlin:2014}
but unidentified by that analysis.

@@

\acknowledgments TDM acknowledges support from the \kepler\
Participating Scientist Program (NNX14AE11G). This paper includes data
collected by the \kepler\ mission. Funding for the \kepler\ mission
is provided by the NASA Science Mission directorate. The authors
acknowledge the efforts of the \kepler\ Mission team for obtaining
the light curve products used in
this publication, which were generated by the \kepler\ Mission
science pipeline through the efforts of the \kepler\ Science
Operations Center and Science Office. The \kepler\ Mission is led by
the project office at NASA Ames Research Center. Ball Aerospace built
the \kepler\ photometer and spacecraft which is operated by the
mission operations center at LASP. These data products are archived at
the NASA Exoplanet Science Institute, which is operated by the
California Institute of Technology, under contract with the National
Aeronautics and Space Administration under the Exoplanet Exploration
Program. This research has made use of NASA's Astrophysics Data
System.

@@

\begin{figure*}[p]
\begin{center}
\includegraphics[width=7in]{figures/fpp_summary_all.pdf}
\end{center}
\caption{False positive probabilities of all CANDIDATE or CONFIRMED
  KOIs for which we consider the \vespa\ calculations to be reliable
  (floored at $10^{-4}$ for visualization purposes), meaning they are
  considered to be reliably located on the target star ($ {\rm Pr} >
  0.99$, with ``score'' $> \posprobthresh$) according to pixel-level analysis
  \citep{Bryson:KSCI}, and have $L_{\rm tot} > 10^{-3}$, indicating
  that the hypotheses tested by \vespa\ are not all poor explanations
  of the data (see \sect{results:fpp}).  Of these \nreliable\ KOIs,
  \nval\ have FPPs less than 1\% (\nvalnew\ of which have not yet been
  dispositioned as CONFIRMED).  Noteably, \nreliableFP\ are likely
  false positives (FPP $> 0.5$), consistent with the
  \citet{Morton:2011b} and \citet{Fressin:2013} \emph{a priori}
  estimates of the overall \kepler\ candidate false positive rate.
  \figlabel{fppall}}
\end{figure*}

@@

\begin{figure*}[p]
\begin{center}
\includegraphics[width=7in]{figures/fpp_pie.pdf}
\end{center}
\caption{A summary of how the the calculations presented in this paper
  advance our understanding of the true nature of KOIs.  More than
  half of all KOIs to date have already been dispositioned FALSE
  POSITIVE or CONFIRMED.  Those dispositioned CANDIDATE we further
  categorize according to their reliability.  The ``no reliable
  calculation'' category means that either the \vespa\ calculation was
  not successful, or that $L_{\rm tot} < 10^{-3}$.  ``On-target
  probability uncertain'' indicates that the positional probability
  calculations of \citet{Bryson:KSCI} are not reliable (score $<
  \posprobthresh$).  ``On-target probability $<$ 0.99'' means that the positional
  probability calculations indicate that there is a non-negligible
  chance that the source of the transit signal is not at the position
  of the KOI.  The remaining three categories are all CANDIDATE KOIs
  reliably confirmed to be located at the presumed target star,
  grouped by false positive probability.  \nvalnew\ of these are new
  planet validations.%, and \nfpnew\ are new likely false positives.
  \figlabel{fpppie}}
\end{figure*}

@@

>>>>

\newcommand{\ntotal}{7470}
\newcommand{\nfail}{414}
\newcommand{\ncalc}{7056}

@@

\newcommand{\nreliable}{2857} %from figure
\newcommand{\nval}{1935} %from figure
\newcommand{\nreliableFP}{419} %from figure
\newcommand{\nvalnew}{1284} %from pie chart
\newcommand{\nfpnew}{428}  % all candidate with FPP < 0.9
\newcommand{\ncurval}{984} % from pie chart
\newcommand{\nhz}{9}

@@

To quantify uncertainty due to the Monte Carlo nature of this procedure, 
\vespa\ is also able to repeat these calculations any desired number of times 
by bootstrap resampling (with replacement) of the simulated populations, 
and recalculating the likelihoods based on the resampled populations.  
This mitigates the chances for rare outliers in a simulation to significantly 
affect the calculated FPP.

@@

Of the \ntotal\ KOIs in the Q1-Q17 DR24 table at the NExScI Exoplanet
Archive, \vespa\ successfully calculates the FPP for \ncalc.  
\Sect{failures} contains detailed explanations of the failure reasons for 
the \nfail\ KOIs for which we do not present \vespa\ results.  
FPPs and their uncertainties are determined as the mean and standard 
deviations of 10 bootstrap 
recalculations of the initially simulated populations for each KOI (as described in 
\sect{methods:fpp}).  These results are listed in \tab{fpp}.  
The median fractional FPP uncertainty 
for KOIs with 0.001 $<$ FPP $<$ 0.1 (within an order of magnitude of the 
validation threshold) is about 12\%, and this distribution 
is shown in \fig{sigmafpp}.

In order to properly interpret these results, it is necessary to
understand the range of applicability of the \vespa\ calculation.
First of all, this method selects between different specific explanations 
for the transit-like signal, and cannot
comment on whether the signal might be caused by stellar variability
or an instrumental false alarm.  Thus, \vespa\ results on low
signal-to-noise ratio (SNR) candidates that are not clearly
transit-like must be viewed with caution.  This being said, the reason 
for including the artificial 
``boxy'' and ``long'' models in the model selection calculation
 (\sect{methods:fpp}) is to flag signals that do not 
fit well with any of the astrophysical eclipse models---in fact, the ``long''
model is preferred ($>50\%$) by \nlongfp\ KOIs that are already dispositioned
as FALSE POSITIVE.

@@

In the Q1-Q17 DR24 KOI table used as the basis for this work, 980
KOIs were identified as PEMs.  Of these, 187 were not identified 
as false positives by any other method.  Only 15 of these 187 survived 
all the quality cuts described in \sect{results:fpp}. And of these 15, 
only 3 have FPP $< 0.01$.  (See \sect{gaidos} for an example of a KOI caused by a 
``column anomaly'' effect that went unidentified by the \kepler\ pipeline but
to which \vespa\ assigned a high FPP.)  Thus, we expect only about 0.3\% (3/980) of 
as-yet unidentified distant-contamination FPs to end
up with FPP $< 0.01$ according to \vespa. As the fraction of false positives
from pixel contamination but unidentified as PEMs among the entire KOI sample is
estimated to be something around 23\% \citep{Coughlin:2014}, we estimate
that there remains a small ($\sim$0.06\%) residual FPP for all KOIs, even 
when the \vespa-calculated FPP is negligibly tiny.

@@

While the FPP below which to claim planet validation is clearly
an arbitrary choice, there is precedent to using ${\rm FPP} <0.01$ as 
the threshold---\citet{Rowe:2014} validated over 800 multi-planet KOIs
using this number, and \citet{Montet:2015} used it to validate a sample
of K2-Campaign 1 planets.  Adopting this same threshold and adding
the 0.06\% residual FPP estimated in \sect{PEMs} to the \vespa-calculated
value, we find that 
of the KOIs with reliable \vespa\ FPPs, \nval\ have ${\rm FPP}
<0.01$, and are thus validated at the 99\% level.  
These KOIs are labeled as such in \tab{fpp}.  These are not all new validations,
however, since a number of them are already CONFIRMED. \Fig{fpppie}
shows a different kind of summary, grouping KOIs by disposition and
splitting up the CANDIDATES according to the \vespa\ results, showing
that \nvalnew\ KOIs are newly validated at the 99\% level.
\Fig{rpcand} shows the radii and periods of the CONFIRMED and 
CANDIDATE KOIs, with the transparency of the points representing
the \vespa-calculated FPP.  \Fig{validatedcompare} compares the temperature
and radii of the stars hosting these planets between the \citetalias{Huber:2014}
and \isochrones\ analysis.

@@

In addition to the confident validations, we identify \nfpnew\ KOIs
that currently have CANDIDATE disposition but for which \vespa\
calculates FPP $>$ 0.9; these KOIs are likely to be false positives.
As \Fig{rpcand} shows, many of these newly identified false positives
have large radii---this is again because of the dispositioning
philosophy adopted by the \kepler\ team, which does not use any cut in
transit depth or inferred ``planet'' size to identify FALSE POSITIVES.
To identify these likely false positives from the \vespa\
calculations, we do not require CANDIDATES to obey the same selections
we used to ensure a clean sample for validation.  This is because even
if a signal has characteristics such that a low \vespa\ FPP would not
be sufficient to validate it, a high \vespa\ FPP is still sufficient
to cast doubt on its planetary nature. As a demonstration of the
ability of \vespa\ to identify false positives, \Fig{fppfps} shows the
FPPs for KOIs that are dispositioned FALSE POSITIVE and have MES $>$
10. The vast majority of these have large FPPs.

@@

In this work, the only astrophysical false positive scenarios we consider are 
eclisping binary stars (EB, HEB, BEB, and the double-period versions
thereof).  Previous work studying \kepler\ false positive rates
\citep[e.g.][]{Fressin:2013} has also considered the  ``blended transiting
planet'' to be a false positive---i.e.,  a fainter companion
star hosting a transiting planet larger than what would be inferred if
it were transiting the target star.  Because we do not consider such
a scenario to be a false positive, the \vespa\ analysis presented here 
does not quantify its probability.  As a result, we are \textit{not}
able to unambiguously determine the radii of the planets we validate---
all the planet radii listed in \tab{fpp} are based on the assumption
that the planet transits the target star and that the target star is 
unblended.  If the target star is actually a member of an unresolved
binary system (as a significant fraction of KOIs undoubtedly are),
then the true planet radius will be larger (significantly larger 
if transiting a fainter companion).  This was indeed the motivation
for \citet{Fressin:2013} to consider the ``blended transiting planet''
as a false positive; part of the goal of that work was to 
compute the planet occurrence rates in different radius bins.  
However, trying to distinguish between single and binary target
star configurations is beyond the scope of this work; we thus follow
the precedent of \citet[][especially Section 5]{Lissauer:2014} by 
acknowledging the potential for substantial radius
uncertainties among the validated planet sample while nevertheless 
defending the validations themselves as robust.

@@

In this work, we have calculated the astrophysical false positive
probability (FPP) for every \kepler\ object of interest (KOI) in the
Q1-Q17 DR24 table, using the publicly available Python module
\vespa, which implements the procedure introduced in
\citet{Morton:2012}, with improvements in stellar parameter modeling
and the inclusion of new ``double-period'' false positive scenarios 
and artificial models to identify KOIs that are not well explained
by any of the astrophysical models.  We have also for the first time
estimated uncertainties in the \vespa\ calculation, through a bootstrap
resampling procedure.

While the assumptions behind this calculation are not necessarily
valid for every KOI (see \sect{results:fpp}), we have identified
\nvalnew\ KOIs that have reliable FPPs of $< 0.01$, resulting in
validation of their planetary nature at the 99\% confidence level,
more than doubling the number of confirmed \kepler\ exoplanets.  Among
this set of newly validated planets are \nhz\ that are consistent with being
in the habitable zones of their host stars.  We also identify \nfpnew\ new likely
false positive KOIs, although we note that some of these may be due to
unidentified or miscorrected transit timing variations.

The reliability of these calculations depends significantly on the
results of \citet{Bryson:KSCI}, which quantify the probability that
the eclipse/transit signal is spatially coincident on the sky with the
presumed target star.  Without confirmation that the transit signal is
not coming from a significantly displaced source, the sky area used as
part of the prior for the false positive scenarios would need to be
significantly larger than the positional uncertainty value assumed in
this work (described in \sect{methods:fpp}).  Additionally, the
blended false positive scenarios \vespa\ considers are assumed to be
fully contained within the photometric aperture; this assumption would
also be broken if the source of the eclipse were significantly
displaced.  We estimate that perhaps 0.06\% of the planets we validate
could be signals coming from significantly displaced sources, similar 
to those identified as period-epoch matches by \citet{Coughlin:2014}
but unidentified by that analysis.

@@

\acknowledgments TDM acknowledges support from the \kepler\
Participating Scientist Program (NNX14AE11G). We thank the anonymous
referee, Daniel Huber, Jack Lissauer, Juna Kollmeier, and David Hogg 
for providing helpful suggestions that improved both the analysis and 
presentation herein.
This paper includes data collected by the \kepler\ mission. Funding for the \kepler\ mission
is provided by the NASA Science Mission directorate. The authors
acknowledge the efforts of the \kepler\ Mission team for obtaining
the light curve products used in
this publication, which were generated by the \kepler\ Mission
science pipeline through the efforts of the \kepler\ Science
Operations Center and Science Office. The \kepler\ Mission is led by
the project office at NASA Ames Research Center. Ball Aerospace built
the \kepler\ photometer and spacecraft which is operated by the
mission operations center at LASP. These data products are archived at
the NASA Exoplanet Science Institute, which is operated by the
California Institute of Technology, under contract with the National
Aeronautics and Space Administration under the Exoplanet Exploration
Program. This research has made use of NASA's Astrophysics Data
System.

@@

\begin{figure*}[p]
\begin{center}
\includegraphics[width=7in]{figures/sigmafpp.pdf}
\end{center}
\caption{Fractional uncertainties for KOIs with FPPs between
  0.001 and 0.1; that is, within an order of magnitude of the 
  validation threshold.  FPP values and uncertainties 
  are determined by the mean and standard deviations 
  of \vespa\ calculations based on 
  10 bootstrap resamplings (see \sect{methods:fpp}) of 
  a single set of simulated populations for each KOI.
  \figlabel{sigmafpp}}
\end{figure*}

\begin{figure*}[p]
\begin{center}
\includegraphics[width=7in]{figures/fpp_summary_all.pdf}
\end{center}
\caption{False positive probabilities of all CANDIDATE or CONFIRMED
  KOIs for which we consider the \vespa\ calculations to be reliable
  (floored at $10^{-4}$ for visualization purposes), meaning they are
  considered to be reliably located on the target star ($ {\rm Pr} >
  0.99$, with ``score'' $> \posprobthresh$) according to pixel-level analysis
  \citep{Bryson:KSCI}, and have MES $>10$.  Of these \nreliable\ KOIs,
  \nval\ have FPPs less than 1\% (\nvalnew\ of which have not yet been
  dispositioned as CONFIRMED).  Noteably, \nreliableFP\ are likely
  false positives (FPP $> 0.5$), consistent with the
  \citet{Morton:2011b} and \citet{Fressin:2013} \emph{a priori}
  estimates of the overall \kepler\ candidate false positive rate.
  Red circles correspond to median FPP values in equal-sized bins. 
  \figlabel{fppall}}
\end{figure*}

@@

\begin{figure*}[p]
\begin{center}
\includegraphics[width=7in]{figures/fpp_pie.pdf}
\end{center}
\caption{A summary of how the the calculations presented in this paper
  advance our understanding of the true nature of KOIs.  More than
  half of all KOIs to date have already been dispositioned FALSE
  POSITIVE or CONFIRMED.  Those dispositioned CANDIDATE we further
  categorize according to their reliability.  The ``no reliable
  calculation'' category means that the \vespa\ calculation was
  not successful.  ``On-target
  probability uncertain'' indicates that the positional probability
  calculations of \citet{Bryson:KSCI} are not reliable (score $<
  \posprobthresh$).  ``On-target probability $<$ 0.99'' means that the positional
  probability calculations indicate that there is a non-negligible
  chance that the source of the transit signal is not at the position
  of the KOI.  The remaining three categories are all CANDIDATE KOIs
  reliably confirmed to be located at the presumed target star,
  grouped by false positive probability.  \nvalnew\ of these are new
  planet validations.%, and \nfpnew\ are new likely false positives.
  \figlabel{fpppie}}
\end{figure*}

@@

\begin{figure*}[p]
\begin{center}
\includegraphics[width=7in]{figures/validated_compare.pdf}
\end{center}
\caption{Comparison between temperatures and radii estimated from the
  \isochrones\ analysis in this work and those from the
  \kepler\ stellar parameters catalog \citepalias{Huber:2014}, for the 
  sample of stars hosting planets validated in this study.  As 
  seen in previous figures, the \isochrones\ analysis tends to 
  over-predict the stellar radii compared to \citetalias{Huber:2014} 
  for the coolest stars.  A small number of stars are also estimated by 
  the photometric analysis of \citetalias{Huber:2014} to be evolved, 
  but not by \isochrones.
  \figlabel{validatedcompare}}
\end{figure*}
